Choosing a Population Genomics Approach

### 2022-10-26 choosing-population-genetics.Rmd
standardizing capitalization and spacing; adding a citation to poolseq-demograpjy table

```{r,echo=FALSE}
text_tbl <- data.frame(
  Goals = c("Pros","Cons","Population structure, genetic diversity","Demography (mig. rates, pop. size through time)",
            "Signatures of selection, GWAS","Phylogenetic inference","Genetic crosses/mapping panels",
            "molecular evolution (accurate low-freq alleles required)"),
  hcWGS = c("provides high quality, high density genotypes; can be used to improve development of a reference genome","Most expensive per sample, requires a reference genome assembly, large computational resources","Great, but cheaper methods work almost as well","Great if incorporating haplotype-based methods ","Great","Computationally challenging, but good at all divergence levels ","Expensive","Best"),
  lcWGS = c("Provides high density genotypes at a reduced cost","Requires a reference, can produce false (+) heterozygotes, sensitive to batch effects, may be inappropriate to use individual SNP calls for some analyses","Good, but limited to popgen analyses based on genotype likelihoods ","Good, esp. with some high coverage samples to help with imputation","Good","Good for tree shapes but not branch lengths, good for extracting organelle sequences and cost-effective primers for species and hybrid ID studies ","Great","Poor"),
 RADseq = c("Generally cheaper per sample, density of genotypes can be tuned to fit question, does not require a reference","Only 1-5% of genome covered which limits studies on adaptive variation, de novo assembly can result in paralogs without quality control, can be hard to capture the same loci between batches","Great, esp. for many samples","Good for methods based on site frequency spectrum (eg moments), but not ideal for methods using extended haplotype or phasing info","Hotly debated, but unless your genome is way too big, WGS is the way to go","good at shallow to medium divergence levels, but need to play with filtering parameters ", "great as it allows for many invididuals","poor" ),
 Poolseq = c("many individuals can be mixed for a low library prep cost, only option for some larval studies","need higher sequencing depth so it can be more expensive than RAD (but less than other options), no info on individual genotypes (duh), requires a reference, requires some replicates to account for batch effects","good for methods that only use allele frequencies, won't work for individual based methods (eg Admixture)","active area of method development, but still not common except in cases of multiple temporal samples (cite) ","good if covering majority of the genome","would work in certain situations","maybe","poor"),
 Target.Capture = c("allows consistent capture of loci between batches, good for sequencing targeted regions across many samples","expensive to start a new project and requires reference of some kind to design baits (unless using method like eecSeq), only gives info on targeted region so can result in ascertainment bias for some analyses","ascertainment bias is likely to skew results, esp. for measures of genetic diversity","likely not appropriate due to ascertainment bias","only works for targeted regions ","an improvement over RAD as there will be less missing data among samples","poor","good if only interested in targeted regions")
  )

@@

* [@Matz2018-vl]: great intro to ecological genomics methods in species with few to no prior genomic resources, accessible to those just getting started.  
* [@Fuentes-Pardo2017-nu]: a VERY thorough review of the different approaches to generating population genomics data, with non-model species in mind. Some of the content about relative cost and available software is a little dated, but otherwise a very valuable reference.   
* [@Lou2021-me]: Great recent primer on low coverage WGS and how it compares to RADseq and Poolseq.

@@

* [@Benjelloun2019-dq]:an empirical comparison of high coverage WGS, low coverage WGS, random variants (eg. from RADseq), exome capture, and commercial SNP chip for mammals (genome size=2.6Gb). For neutral genomic diversity, found 5k-10k random variants were enough to accurately estimate, but commercial panels/exome capture had ascertainment bias. To detect selection and accurate estimates of LD, at least 1M variants were required. For the studied species, 5x coverage was just as accurate as 12x coverage. Low coverage WGS (< 5x) should use genotype likelihood methods, however, the rate of false positives for heterozygotes is higher than classic genotyping methods at higher coverage.
* [@Dorant2019-fb]:  A case study comparing Poolseq vs. reduced representation seq (GBS) vs Rapture (targeted capture of RAD loci) for inferring population structure with weak genetic differentiation. All three detected the same weak pop structure pattern, but Poolseq gave Fst values 3-5x higher than Rapture/GBS.

>>>>

```{r,echo=FALSE}
text_tbl <- data.frame(
  Goals = c("pros","cons","population structure, genetic diversity","demography (mig. rates, pop. size through time)",
            "signatures of selection, GWAS","phylogenetic inference","genetic crosses/mapping panels",
            "molecular evolution (accurate low-freq alleles required)"),
  hcWGS = c("provides high quality, high density genotypes; can be used to improve development of a reference genome","most expensive per sample, requires a reference genome assembly, large computational resources","great, but cheaper methods work almost as well","great if incorporating haplotype-based methods ","great","computationally challenging, but good at all divergence levels ","expensive","best"),
  lcWGS = c("provides high density genotypes at a reduced cost","requires a reference, can produce false (+) heterozygotes, sensitive to batch effects, may be inappropriate to use individual SNP calls for some analyses","good, but limited to popgen analyses based on genotype likelihoods ","good, esp. with some high coverage samples to help with imputation","good","good for tree shapes but not branch lengths, good for extracting organelle sequences and cost-effective primers for species and hybrid ID studies ","great","poor"),
 RADseq = c("generally cheaper per sample, density of genotypes can be tuned to fit question, does not require a reference","only 1-5% of genome covered which limits studies on adaptive variation, de novo assembly can result in paralogs without quality control, can be hard to capture the same loci between batches","great, esp. for many samples","good for methods based on site frequency spectrum (eg moments), but not ideal for methods using extended haplotype or phasing info","hotly debated, but unless your genome is way too big, WGS is the way to go","good at shallow to medium divergence levels, but need to play with filtering parameters ", "great as it allows for many invididuals","poor" ),
 Poolseq = c("many individuals can be mixed for a low library prep cost, only option for some larval studies","need higher sequencing depth so it can be more expensive than RAD (but less than other options), no info on individual genotypes (duh), requires a reference, requires some replicates to account for batch effects","good for methods that only use allele frequencies, won't work for individual based methods (eg Admixture)","active area of method development, but still not common except in cases of multiple temporal samples [@nielsen2020bc] ","good if covering majority of the genome","would work in certain situations","maybe","poor"),
 Target.Capture = c("allows consistent capture of loci between batches, good for sequencing targeted regions across many samples","expensive to start a new project and requires reference of some kind to design baits (unless using method like eecSeq), only gives info on targeted region so can result in ascertainment bias for some analyses","ascertainment bias is likely to skew results, esp. for measures of genetic diversity","likely not appropriate due to ascertainment bias","only works for targeted regions ","an improvement over RAD as there will be less missing data among samples","poor","good if only interested in targeted regions")
  )

@@

* [@Matz2018-vl]: great intro to ecological genomics methods in species with few to no prior genomic resources, accessible to those just getting started.  
* [@Fuentes-Pardo2017-nu]: a VERY thorough review of the different approaches to generating population genomics data, with non-model species in mind. Some of the content about relative cost and available software is a little dated, but otherwise a very valuable reference.   
* [@Lou2021-me]: great recent primer on low coverage WGS and how it compares to RADseq and Poolseq.

@@

* [@Benjelloun2019-dq]: an empirical comparison of high coverage WGS, low coverage WGS, random variants (eg. from RADseq), exome capture, and commercial SNP chip for mammals (genome size=2.6Gb). For neutral genomic diversity, found 5k-10k random variants were enough to accurately estimate, but commercial panels/exome capture had ascertainment bias. To detect selection and accurate estimates of LD, at least 1M variants were required. For the studied species, 5x coverage was just as accurate as 12x coverage. Low coverage WGS (< 5x) should use genotype likelihood methods, however, the rate of false positives for heterozygotes is higher than classic genotyping methods at higher coverage.
* [@Dorant2019-fb]: a case study comparing Poolseq vs. reduced representation seq (GBS) vs Rapture (targeted capture of RAD loci) for inferring population structure with weak genetic differentiation. All three detected the same weak pop structure pattern, but Poolseq gave Fst values 3-5x higher than Rapture/GBS.
